---
output:
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
    toc_depth: 3
---




<!-- Javascript for zooming on figures (adapted from: https://stackoverflow.com/questions/40401680) -->

<!-- Jquery import conflicts with DT::datatable so needs to be commented here -->
<!-- <script src="https://ajax.googleapis.com/ajax/libs/jquery/3.1.1/jquery.min.js"></script> -->

<style>
.zoomDiv {
  display: none;
  position: fixed;
  top: 50%;
  left: 50%;
  z-index: 50;
  transform: translate(-50%, -50%);
  background-color: #FFFFFF;
  box-shadow: 0px 0px 50px #888888;
  width: fit-content;
  max-width: 90%;
  max-height: 90%;
  overflow: auto;
}

.zoomImg {
  width: 100%;
}
</style>

<script type="text/javascript">
  $(document).ready(function() {
    $('body').prepend("<div class=\"zoomDiv\"><img src=\"\" class=\"zoomImg\"></div>");
    // onClick for all img except the zoomed one and link ones (filter)
    // use 'img.zoom' and out.extra='class=\"zoom\"' in chunk to specify manually which chunk images can be zoomed
    $('img:not(.zoomImg)').filter(':not(a *)').click(function() {
      $('.zoomImg').attr('src', $(this).attr('src'));
      $('.zoomDiv').show();
    });
    // onClick function for hiding div
    $('img.zoomImg').click(function() {
      $('.zoomDiv').hide();
    });
  });
</script>




<!-- Load spectrum color picker code -->
```{js loadSpectrumJS, echo = FALSE, code=readLines("spectrum.js")}
```
```{css loadSpectrumCSS, echo = FALSE, code=readLines("spectrum.css")}
```

<!-- Prepare a js function extracting traces color from plotly figure -->
```{js getColorTab, echo = FALSE}

// Ask figure template from plotly, extract markers color (scatter and scattergl only), and place them in clipboard
getColorTab = function(plotLyFigure, type = "scatter", colorFormatFun = "hex") 
  { 
    var template = Plotly.makeTemplate(plotLyFigure);
    var resultColorString = "";
    for (curTrace in template.data[type])
    {
      // Only use traces having markers
      if(template.data[type][curTrace].mode == "markers") resultColorString += '"' + template.data[type][curTrace].name + '" = "' + tinycolor(template.data[type][curTrace].marker.color).toString(colorFormatFun) + '"\n';
    }
    // Create a dummy element to copy content to clipboard
    var textarea = document.createElement("textarea"); textarea.textContent = resultColorString; textarea.style.position = "fixed"; document.body.appendChild(textarea); textarea.select(); document.execCommand("copy"); document.body.removeChild(textarea);
    alert("\n" + resultColorString + "\n\nCopied to clipboard !");
  } 
```




```{r rmd_setKnitrOptions, include=FALSE}
# Set hook for PNG optimization ('optipng' must be installed on system)
#knitr::knit_hooks$set(optipng = knitr::hook_optipng)

# Set chunks default options
knitr::opts_chunk$set(results='asis',
                      echo = FALSE,           # Do not show chunks code
                      message = FALSE,        # Hide R regular text output in reports
                      warning = TRUE,         # Show warning messages
                      error = FALSE,          # Stop generating report on errors
                      cache = FALSE,          # Disable chunk result caching
                      cache.comments = FALSE, # Changing comments does not invalidate cache
#                      dev = "svg",            # Generate all figures as vectorized 'svg' files                      
                      fig.dim = c(8, 8),      # Size of device for R figures
                      autodep = TRUE,         # Automatic chunk dependencies if cache enabled
                      optipng = '-o2');       # PNG optimization level (2=default, 5=slow, 7=very slow)
```




```{r rmd_loadParamsAndDependencies, warning=FALSE, cache = FALSE}
### Define working folder
WORKING_DIR   = getwd();


### Load libraries, global functions/variables, and chunks
source( file.path( WORKING_DIR, "00_generalDeps.R"));

# Import chunks from source files
read_chunk( path=file.path( WORKING_DIR, "01_prepareData.R"));
read_chunk( path=file.path( WORKING_DIR, "02_heatmaps.R"));
read_chunk( path=file.path( WORKING_DIR, "03_moduleScore.R"));
read_chunk( path=file.path( WORKING_DIR, "04_cellsCounts.R"));
read_chunk( path=file.path( WORKING_DIR, "90_finalize.R"));
```




---
title: `r paste( GLOBAL_DESCRIPTION, ANALYSIS_STEP_NAME, sep ="<BR>")`
---




```{r rmd_initFunctions}

### Create output directory
dir.create( PATH_ANALYSIS_OUTPUT, showWarnings = FALSE);


### Record starting time (for report and file names)
startTime = Sys.time();
startTimeReport   = format( startTime, "%Y %m %d %H:%M:%S")
startTimeFileName = format( startTime, "%Y_%m_%d_%H_%M_%S")


### Initialize utils

# Create a hash for all parameters
paramsHash = digest( environment(), algo="murmur32");

# Create a 'safe' unique prefix for output files (now specified at project scale in globalParams.R)
#outputFilesPrefix = path_sanitize( gsub("([[:punct:]])|\\s+",
#                                        "_",
#                                        paste0(PROJECT_NAME, "_",
#                                               SAMPLE_NAME, "_",
#                                               startTimeFileName, "_",
#                                               paramsHash, "_")),
#                                   replacement = "_");

# Set seed for pseudo random number generator
set.seed(SEED);

# Set the use of multiple cores using 'future' library, as implemented in Seurat3 when possible
# https://github.com/satijalab/seurat/issues/457
plan(strategy = "multicore", workers = NBCORES);

# Create a hidden plotly in a tagList to allow using them in loops later
cat('<div style="display: none">');
htmltools::tagList(plot_ly(x = rnorm(10), type = "histogram"))
cat('</div>');

# Same for datatable....
cat('<div style="display: none">');
datatable( data.frame(a=1:10, b = 1:10),
           class = "compact",
           filter="top",
           rownames = FALSE,
           #colnames = ,
           #caption = ,
           extensions = c('Buttons', 'Select'))
cat('</div>');
```




## Run informations {.tabset .tabset-fade}

### Run
<center>
```{r rmd_runInfo_run}
pander( c( 'Date'     = startTimeReport,
           'Sample'   = EXPERIMENT_PROJECT_NAME,
           'Hash'     = paramsHash));
```
</center>

### Parameters
```{r rmd_runInfo_params}
listParams = as.list( environment());
pander(listParams[order(names(listParams))]); # Show params in alphabetical order
```

### System
<center>
```{r rmd_runInfo_sysInfo}
showSimpleDT( Sys.info())
```
</center>

### R
```{r rmd_runInfo_r}
pander( sessionInfo());
```

<!-- Ending the tabset section -->
## {.toc-ignore}
<br /><br />




## Load data
```{r rmd_loadData}
# Check whether results from previous run are available, in which case this is 
# a second run for clean rendering (generation of external figures interferes 
# with rmarkdown layout on first run). See analysisParams.
sessionImagePath = file.path( PATH_ANALYSIS_OUTPUT, paste0( outputFilesPrefix, "sessionImage_final.RDATA"));
FLAG_secondRun = file.exists( sessionImagePath) # FLAG specifying if this is a second run (time consuming steps can be skipped)
if(FLAG_secondRun) 
{
  # Do not recompute, just load saved env and integrate external png plots in current report
  load( sessionImagePath)
  FLAG_secondRun = TRUE
  warning("Second run (for rendering), results loaded from first run. Tabset 'parameters' may not reflect values used for analysis.")
  print(sc10x)
} else
{
<<loadData>>
<<loadGenesLists>>
<<filtering>>
}

```
<br /><br />  



## Genes lists

### Heatmaps by groups of cells {.tabset .tabset-fade}

```{r rmd_genes_heatmaps}
<<heatmap_compute_expression_summary_matrices>>
<<heatmap_function_definition>>
<<heatmaps_genes>>
```
<br /><br />  



### Expression bubbles {.tabset .tabset-fade}
```{r rmd_expression_bubbles_eachCluster_byHTO, fig.dim = c(4, 10)}
<<expression_bubbles_eachCluster_byHTO>>
```
<br /><br />  



### Expression violin by HTO {.tabset .tabset-fade}
```{r rmd_cellsCounts_violin_eachCluster_byHTO, fig.dim = c(6, 15)}
<<cellsCounts_violin_eachCluster_byHTO>>
```
<br /><br />  



### Expression violin by cluster {.tabset .tabset-fade}
```{r rmd_cellsCounts_violin, fig.dim = c(6, 15)}
<<cellsCounts_violin>>
```




## Modules score
```{r rmd_heterogeneity_modules}
<<heterogeneity_modules_scoring>>
```

### Heatmap

Heatmap of modules scoring sorted by Seurat clusters:

```{r rmd_heterogeneity_modules_heatmap, fig.dim = c(10, 5)}
<<heterogeneity_modules_heatmap>>
```

### Modules scoring {.tabset .tabset-fade}

#### Violin by HTO {.tabset .tabset-fade}
```{r rmd_heterogeneity_modules_expression_violin_byHTO, fig.dim = c(4, 4), out.width='25%'}
cat("\n \n");
<<heterogeneity_modules_expression_violin_byHTO>>
cat("\n \n");
```

#### Clusters violin {.tabset .tabset-fade}
```{r rmd_heterogeneity_modules_expression_violin, fig.dim = c(4, 4), out.width='25%'}
cat("\n<a href='#clusters-projection'>");
useReduction='umap'; #c( 'umap', 'tsne');
<<heterogeneity_dimReduc_thumbnail>>
rm("useReduction");
cat("</a> \n \n");

<<heterogeneity_modules_expression_violin>>
<<heterogeneity_modules_expression_violin_pngFile>>
cat("\n \n");

```

#### UMAP {.tabset .tabset-fade .tabset-pills}
```{r rmd_heterogeneity_modules_expression_projection_UMAP, fig.dim = c(4, 4), out.width='25%'}
useReduction='umap'

cat("\n<a href='#clusters-projection'>");
<<heterogeneity_dimReduc_thumbnail>>
cat("</a> \n \n");

<<heterogeneity_modules_expression_projection>>
<<heterogeneity_modules_expression_projection_pngFile>>

rm("useReduction");
```



<!-- Ending the tabset section -->
#### {.toc-ignore}



### Cells population {.tabset .tabset-fade}
```{r rmd_cellsCounts, fig.dim = c(10.5, 7)}
<<cellsCounts>>
```




### Combinatorial expression {.tabset .tabset-fade}
```{r rmd_cellsCounts_genesCombinations, warning=FALSE, fig.dim = c(9, 5)}
<<cellsCounts_genesCombinations>>
```




```{r rmd_final_saveSessionImage}
if(!FLAG_secondRun) # check if this is a second run for rendering only (see analysisParams)
{
<<final_saveSessionImage>>
}
```


