---
output:
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
    toc_depth: 3
---




<!-- Javascript for zooming on figures (adapted from: https://stackoverflow.com/questions/40401680) -->

<!-- Jquery import conflicts with DT::datatable so needs to be commented here -->
<!-- <script src="https://ajax.googleapis.com/ajax/libs/jquery/3.1.1/jquery.min.js"></script> -->

<style>
.zoomDiv {
  display: none;
  position: fixed;
  top: 50%;
  left: 50%;
  z-index: 50;
  transform: translate(-50%, -50%);
  background-color: #FFFFFF;
  box-shadow: 0px 0px 50px #888888;
  width: fit-content;
  max-width: 90%;
  max-height: 90%;
  overflow: auto;
}

.zoomImg {
  width: 100%;
}
</style>

<script type="text/javascript">
  $(document).ready(function() {
    $('body').prepend("<div class=\"zoomDiv\"><img src=\"\" class=\"zoomImg\"></div>");
    // onClick for all img except the zoomed one and link ones (filter)
    // use 'img.zoom' and out.extra='class=\"zoom\"' in chunk to specify manually which chunk images can be zoomed
    $('img:not(.zoomImg)').filter(':not(a *)').click(function() {
      $('.zoomImg').attr('src', $(this).attr('src'));
      $('.zoomDiv').show();
    });
    // onClick function for hiding div
    $('img.zoomImg').click(function() {
      $('.zoomDiv').hide();
    });
  });
</script>




<!-- Load spectrum color picker code -->
```{js loadSpectrumJS, echo = FALSE, code=readLines("spectrum.js")}
```
```{css loadSpectrumCSS, echo = FALSE, code=readLines("spectrum.css")}
```

<!-- Prepare a js function extracting traces color from plotly figure -->
```{js getColorTab, echo = FALSE}

// Ask figure template from plotly, extract markers color (scatter and scattergl only), and place them in clipboard
getColorTab = function(plotLyFigure, type = "scatter", colorFormatFun = "hex") 
  { 
    var template = Plotly.makeTemplate(plotLyFigure);
    var resultColorString = "";
    for (curTrace in template.data[type])
    {
      // Only use traces having markers
      if(template.data[type][curTrace].mode == "markers") resultColorString += '"' + template.data[type][curTrace].name + '" = "' + tinycolor(template.data[type][curTrace].marker.color).toString(colorFormatFun) + '"\n';
    }
    // Create a dummy element to copy content to clipboard
    var textarea = document.createElement("textarea"); textarea.textContent = resultColorString; textarea.style.position = "fixed"; document.body.appendChild(textarea); textarea.select(); document.execCommand("copy"); document.body.removeChild(textarea);
    alert("\n" + resultColorString + "\n\nCopied to clipboard !");
  } 
```




```{r rmd_setKnitrOptions, include=FALSE}
# Set hook for PNG optimization ('optipng' must be installed on system)
#knitr::knit_hooks$set(optipng = knitr::hook_optipng)

# Set chunks default options
knitr::opts_chunk$set(results='asis',
                      echo = FALSE,           # Do not show chunks code
                      message = FALSE,        # Hide R regular text output in reports
                      warning = TRUE,         # Show warning messages
                      error = FALSE,          # Stop generating report on errors
                      cache = FALSE,          # Disable chunk result caching
                      cache.comments = FALSE, # Changing comments does not invalidate cache
#                      dev = "svg",            # Generate all figures as vectorized 'svg' files                      
                      fig.dim = c(8, 8),      # Size of device for R figures
                      autodep = TRUE,         # Automatic chunk dependencies if cache enabled
                      optipng = '-o2');       # PNG optimization level (2=default, 5=slow, 7=very slow)
```




```{r rmd_loadParamsAndDependencies, warning=FALSE, cache = FALSE}
### Define working folder
WORKING_DIR   = getwd();


### Load libraries, global functions/variables, and chunks
source( file.path( WORKING_DIR, "00_generalDeps.R"));

# Import chunks from source files
read_chunk( path=file.path( WORKING_DIR, "01_prepareData.R"));
read_chunk( path=file.path( WORKING_DIR, "02_plots.R"));
read_chunk( path=file.path( WORKING_DIR, "03_diffExp_HTOs.R"));
read_chunk( path=file.path( WORKING_DIR, "90_finalize.R"));
```




---
title: `r paste( GLOBAL_DESCRIPTION, ANALYSIS_STEP_NAME, sep ="<BR>")`
---




```{r rmd_initFunctions}

### Create output directory
dir.create( PATH_ANALYSIS_OUTPUT, showWarnings = FALSE);


### Record starting time (for report and file names)
startTime = Sys.time();
startTimeReport   = format( startTime, "%Y %m %d %H:%M:%S")
startTimeFileName = format( startTime, "%Y_%m_%d_%H_%M_%S")


### Initialize utils

# Create a hash for all parameters
paramsHash = digest( environment(), algo="murmur32");

# Create a 'safe' unique prefix for output files (now specified at project scale in globalParams.R)
#outputFilesPrefix = path_sanitize( gsub("([[:punct:]])|\\s+",
#                                        "_",
#                                        paste0(PROJECT_NAME, "_",
#                                               SAMPLE_NAME, "_",
#                                               startTimeFileName, "_",
#                                               paramsHash, "_")),
#                                   replacement = "_");

# Set seed for pseudo random number generator
set.seed(SEED);

# Set the use of multiple cores using 'future' library, as implemented in Seurat3 when possible
# https://github.com/satijalab/seurat/issues/457
plan(strategy = "multicore", workers = NBCORES);

# Create a hidden plotly in a tagList to allow using them in loops later
cat('<div style="display: none">');
htmltools::tagList(plot_ly(x = rnorm(10), type = "histogram"))
cat('</div>');

# Same for datatable....
cat('<div style="display: none">');
datatable( data.frame(a=1:10, b = 1:10),
           class = "compact",
           filter="top",
           rownames = FALSE,
           #colnames = ,
           #caption = ,
           extensions = c('Buttons', 'Select'))
cat('</div>');
```




## Run informations {.tabset .tabset-fade}

### Run
<center>
```{r rmd_runInfo_run}
pander( c( 'Date'     = startTimeReport,
           'Sample'   = EXPERIMENT_PROJECT_NAME,
           'Hash'     = paramsHash));
```
</center>

### Parameters
```{r rmd_runInfo_params}
listParams = as.list( environment());
pander(listParams[order(names(listParams))]); # Show params in alphabetical order
```

### System
<center>
```{r rmd_runInfo_sysInfo}
showSimpleDT( Sys.info())
```
</center>

### R
```{r rmd_runInfo_r}
pander( sessionInfo());
```

<!-- Ending the tabset section -->
## {.toc-ignore}
<br /><br />




## Load data
```{r rmd_loadData}
# Check whether results from previous run are available, in which case this is 
# a second run for clean rendering (generation of external figures interferes 
# with rmarkdown layout on first run). See analysisParams.
sessionImagePath = file.path( PATH_ANALYSIS_OUTPUT, paste0( outputFilesPrefix, "sessionImage_final.RDATA"));
FLAG_secondRun = file.exists( sessionImagePath) # FLAG specifying if this is a second run (time consuming steps can be skipped)
if(FLAG_secondRun) 
{
  load( sessionImagePath)
  FLAG_secondRun = TRUE
  warning("Second run (for rendering), results loaded from first run. Tabset 'parameters' may not reflect values used for analysis.")
  print(sc10x)
} else
{
<<loadData>>
}
<<plotsPreparation>>
```
<br /><br />  




## Clusters size
```{r rmd_plotClustersSize}
<<plotClustersSize>>
```

### HTOs vs Clusters counts analysis (X²)

X² test of independance between samples/HTOs and identified cell populations/clusters.
```{r rmd_chisquare_test_clusters_HTO}
<<chisquare_test_clusters_HTO>>
```
<br />  

#### Test statistics

Pearson residuals, and contribution (%) to total X² statistic (100*(residuals^2)/statistic):
```{r rmd_chisquare_test_clusters_HTO_plotResidualsAndContributions, fig.dim = c(10.5, 7)}
<<chisquare_test_clusters_HTO_plotResidualsAndContributions>>
```
<br />  

Mosaic-plot (surface of an element reflects the relative magnitude of its value):
```{r rmd_chisquare_test_clusters_HTO_plotMosaic, fig.dim = c(8, 7)}
<<chisquare_test_clusters_HTO_plotMosaic>>
```
<br />  




## Dimensionality reduction {.tabset .tabset-fade}

### Clusters {.tabset .tabset-fade .tabset-pills}
<!-- #### Interactive -->
<!-- ```{r rmd_plotDimReducInteractive_colorClusters} -->
<!-- dimReducWidth  = 950; -->
<!-- dimReducHeight = 800; -->
<!-- initialPointSize = 5; -->
<!-- showDimReducClusterLabels = FALSE; -->
<!-- showLegend = TRUE; -->
<!-- <<plotDimReducInteractive_colorClusters>> -->
<!-- rm(showDimReducClusterLabels); -->
<!-- rm(showLegend); -->
<!-- ``` -->
<!-- <br />   -->

<!-- #### Raster -->
```{r rmd_plotDimReducRaster_colorClusters}
showDimReducClusterLabels = TRUE;
showLegend = FALSE;
<<plotDimReducRaster_colorClusters>>
rm(showDimReducClusterLabels);
rm(showLegend);
```
<br />  

### Clusters + HTOs {.tabset .tabset-fade .tabset-pills}
<!-- #### Interactive -->
<!-- ```{r rmd_plotDimReducInteractive_colorClusters_symbolHTOs} -->
<!-- dimReducWidth  = 950; -->
<!-- dimReducHeight = 800; -->
<!-- initialPointSize = 5; -->
<!-- showDimReducClusterLabels = FALSE; -->
<!-- showLegend = TRUE; -->
<!-- <<plotDimReducInteractive_colorClusters_symbolHTOs>> -->
<!-- rm(showDimReducClusterLabels); -->
<!-- rm(showLegend); -->
<!-- ``` -->
<!-- <br />   -->

<!-- #### Raster -->
```{r rmd_plotDimReducRaster_colorClusters_symbolHTOs, fig.dim = c(11, 9)}
showDimReducClusterLabels = FALSE;
showLegend = TRUE;
<<plotDimReducRaster_colorClusters_symbolHTOs>>
rm(showDimReducClusterLabels);
rm(showLegend);
```
<br />  

### Facet HTOs {.tabset .tabset-fade .tabset-pills}
<!-- #### Interactive -->
<!-- ```{r rmd_plotDimReducInteractiveGG_colorClusters_facetHTOs, warning = FALSE} -->
<!-- dimReducWidth  = 950; -->
<!-- dimReducHeight = 800; -->
<!-- initialPointSize = 5; -->
<!-- showDimReducClusterLabels = FALSE; -->
<!-- showLegend = TRUE; -->
<!-- <<plotDimReducInteractiveGG_colorClusters_facetHTOs>> -->
<!-- rm(showDimReducClusterLabels); -->
<!-- rm(showLegend); -->
<!-- ``` -->
<!-- <br />   -->

<!-- #### Raster -->
```{r rmd_plotDimReducRaster_colorClusters_facetHTOs, fig.dim = c(11, 9)}
showDimReducClusterLabels = FALSE;
showLegend = TRUE;
<<plotDimReducRaster_colorClusters_facetHTOs>>
rm(showDimReducClusterLabels);
rm(showLegend);
```
<br />  

### Facet Clusters {.tabset .tabset-fade .tabset-pills}
<!-- #### Interactive -->
<!-- ```{r rmd_plotDimReducInteractiveGG_colorHTOs_facetClusters, warning = FALSE} -->
<!-- dimReducWidth  = 950; -->
<!-- dimReducHeight = 800; -->
<!-- initialPointSize = 5; -->
<!-- showDimReducClusterLabels = FALSE; -->
<!-- showLegend = TRUE; -->
<!-- <<plotDimReducInteractiveGG_colorHTOs_facetClusters>> -->
<!-- rm(showDimReducClusterLabels); -->
<!-- rm(showLegend); -->
<!-- ``` -->
<!-- <br />   -->

<!-- #### Raster -->
```{r rmd_plotDimReducRaster_colorHTOs_facetClusters_allInBackground, fig.dim = c(11, 9)}
showDimReducClusterLabels = FALSE;
showLegend = TRUE;
<<plotDimReducRaster_colorHTOs_facetClusters_allInBackground>>
rm(showDimReducClusterLabels);
rm(showLegend);
```
<br />  




### Density: ref groups
```{r rmd_plotDimReducRaster_colorDensity_circleDensity_globalRefGroups, warning=FALSE, fig.dim = c(11, 9)}
<<plotDimReducRaster_colorDensity_circleDensity_prepareData>>

<<plotDimReducRaster_colorDensity_circleDensity_globalRefGroups>>
```
<br />  

### Density: Facet HTOs
```{r rmd_plotDimReducRaster_colorDensity_circleDensity_facetByHTO, warning=FALSE, fig.dim = c(9, 9)}
<<plotDimReducRaster_colorDensity_circleDensity_facetByHTO>>
```
<br />  

### Density: Facet Clusters + HTOs
```{r rmd_plotDimReducRaster_colorDensity_circleDensity_facetByHTOAndCluster, warning=FALSE, fig.dim = c(0.5+(2.5*length(table(cellsData[["Cluster"]]))), 0.5+(2.25*length(HTO_FACTOR_LEVELS)))}
<<plotDimReducRaster_colorDensity_circleDensity_facetByHTOAndCluster>>
```
<br />  

### Density: Facet Clusters + HTOs grouped
```{r rmd_plotDimReducRaster_colorDensity_circleDensity_facetByHTOGroupedAndCluster, warning=FALSE, fig.dim = c(0.5+(2.7*length(table(cellsData[["Cluster"]]))), 5)}
<<plotDimReducRaster_colorDensity_circleDensity_facetByHTOGroupedAndCluster>>
```
<br />  



















## Differential expression {.tabset .tabset-fade}

```{r rmd_diffExp_computeDE}
if(!FLAG_secondRun) # check if this is a second run for rendering only (see analysisParams)
{
<<diffExp_computeDE>>
# Set universe to the global list of annotations (instead of union of DE genes)
UNIVERSE_IS_UNIONGENES = FALSE;
<<functional_enrichment_preparation>>
<<functional_kegg_pathway>>
<<functional_kegg_module>>
<<functional_go>>
}
```


```{r rmd_diffExp_plotDE, warning = FALSE, fig.dim = c( 10, 10)}
# Ignore warnings as it prevents proper ordering of figures in tabsets...
# Warning for empty enrichments raised in previous chunks
# Has internal check for 'FLAG_secondRun' to know whether it should generate external figures or integrate previously generated ones
<<diffExp_plots>>
```



```{r rmd_final_saveSessionImage}
if(!FLAG_secondRun) # check if this is a second run for rendering only (see analysisParams)
{
<<final_saveSessionImage>>
}
```



