---
output:
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
    toc_depth: 3
---




<!-- Javascript for zooming on figures (adapted from: https://stackoverflow.com/questions/40401680) -->

<!-- Jquery import conflicts with DT::datatable so needs to be commented here -->
<!-- <script src="https://ajax.googleapis.com/ajax/libs/jquery/3.1.1/jquery.min.js"></script> -->

<style>
.zoomDiv {
  display: none;
  position: fixed;
  top: 50%;
  left: 50%;
  z-index: 50;
  transform: translate(-50%, -50%);
  background-color: #FFFFFF;
  box-shadow: 0px 0px 50px #888888;
  width: fit-content;
  max-width: 90%;
  max-height: 90%;
  overflow: auto;
}

.zoomImg {
  width: 100%;
}
</style>

<script type="text/javascript">
  $(document).ready(function() {
    $('body').prepend("<div class=\"zoomDiv\"><img src=\"\" class=\"zoomImg\"></div>");
    // onClick for all img except the zoomed one and link ones (filter)
    // use 'img.zoom' and out.extra='class=\"zoom\"' in chunk to specify manually which chunk images can be zoomed
    $('img:not(.zoomImg)').filter(':not(a *)').click(function() {
      $('.zoomImg').attr('src', $(this).attr('src'));
      $('.zoomDiv').show();
    });
    // onClick function for hiding div
    $('img.zoomImg').click(function() {
      $('.zoomDiv').hide();
    });
  });
</script>




<!-- Load spectrum color picker code -->
```{js loadSpectrumJS, echo = FALSE, code=readLines("spectrum.js")}
```
```{css loadSpectrumCSS, echo = FALSE, code=readLines("spectrum.css")}
```

<!-- Prepare a js function extracting traces color from plotly figure -->
```{js getColorTab, echo = FALSE}

// Ask figure template from plotly, extract markers color (scatter and scattergl only), and place them in clipboard
getColorTab = function(plotLyFigure, type = "scatter", colorFormatFun = "hex") 
  { 
    var template = Plotly.makeTemplate(plotLyFigure);
    var resultColorString = "";
    for (curTrace in template.data[type])
    {
      // Only use traces having markers
      if(template.data[type][curTrace].mode == "markers") resultColorString += '"' + template.data[type][curTrace].name + '" = "' + tinycolor(template.data[type][curTrace].marker.color).toString(colorFormatFun) + '"\n';
    }
    // Create a dummy element to copy content to clipboard
    var textarea = document.createElement("textarea"); textarea.textContent = resultColorString; textarea.style.position = "fixed"; document.body.appendChild(textarea); textarea.select(); document.execCommand("copy"); document.body.removeChild(textarea);
    alert("\n" + resultColorString + "\n\nCopied to clipboard !");
  } 
```




```{r rmd_setKnitrOptions, include=FALSE}
# Set hook for PNG optimization ('optipng' must be installed on system)
#knitr::knit_hooks$set(optipng = knitr::hook_optipng)

# Set chunks default options
knitr::opts_chunk$set(results='asis',
                      echo = FALSE,           # Do not show chunks code
                      message = FALSE,        # Hide R regular text output in reports
                      warning = TRUE,         # Show warning messages
                      error = FALSE,          # Stop generating report on errors
                      cache = FALSE,          # Disable chunk result caching
                      cache.comments = FALSE, # Changing comments does not invalidate cache
#                      dev = "svg",            # Generate all figures as vectorized 'svg' files                      
                      fig.dim = c(8, 8),      # Size of device for R figures
                      autodep = TRUE,         # Automatic chunk dependencies if cache enabled
                      optipng = '-o2');       # PNG optimization level (2=default, 5=slow, 7=very slow)
```




```{r rmd_loadParamsAndDependencies, warning=FALSE, cache = FALSE}
### Define working folder
WORKING_DIR   = getwd();


### Load libraries, global functions/variables, and chunks
source( file.path( WORKING_DIR, "00_generalDeps.R"));

# Import chunks from source files
read_chunk( path=file.path( WORKING_DIR, "01_prepareData.R"));
read_chunk( path=file.path( WORKING_DIR, "02_plots.R"));
read_chunk( path=file.path( WORKING_DIR, "03_diffExp_HTOs.R"));
read_chunk( path=file.path( WORKING_DIR, "90_finalize.R"));
```




---
title: `r paste( GLOBAL_DESCRIPTION, ANALYSIS_STEP_NAME, sep ="<BR>")`
---




```{r rmd_initFunctions}

### Create output directory
dir.create( PATH_ANALYSIS_OUTPUT, showWarnings = FALSE);


### Record starting time (for report and file names)
startTime = Sys.time();
startTimeReport   = format( startTime, "%Y %m %d %H:%M:%S")
startTimeFileName = format( startTime, "%Y_%m_%d_%H_%M_%S")


### Initialize utils

# Create a hash for all parameters
paramsHash = digest( environment(), algo="murmur32");

# Create a 'safe' unique prefix for output files (now specified at project scale in globalParams.R)
#outputFilesPrefix = path_sanitize( gsub("([[:punct:]])|\\s+",
#                                        "_",
#                                        paste0(PROJECT_NAME, "_",
#                                               SAMPLE_NAME, "_",
#                                               startTimeFileName, "_",
#                                               paramsHash, "_")),
#                                   replacement = "_");

# Set seed for pseudo random number generator
set.seed(SEED);

# Set the use of multiple cores using 'future' library, as implemented in Seurat3 when possible
# https://github.com/satijalab/seurat/issues/457
plan(strategy = "multicore", workers = NBCORES);

# Create a hidden plotly in a tagList to allow using them in loops later
cat('<div style="display: none">');
htmltools::tagList(plot_ly(x = rnorm(10), type = "histogram"))
cat('</div>');

# Same for datatable....
cat('<div style="display: none">');
datatable( data.frame(a=1:10, b = 1:10),
           class = "compact",
           filter="top",
           rownames = FALSE,
           #colnames = ,
           #caption = ,
           extensions = c('Buttons', 'Select'))
cat('</div>');
```




## Run informations {.tabset .tabset-fade}

### Run
<center>
```{r rmd_runInfo_run}
pander( c( 'Date'     = startTimeReport,
           'Sample'   = EXPERIMENT_PROJECT_NAME,
           'Hash'     = paramsHash));
```
</center>

### Parameters
```{r rmd_runInfo_params}
listParams = as.list( environment());
pander(listParams[order(names(listParams))]); # Show params in alphabetical order
```

### System
<center>
```{r rmd_runInfo_sysInfo}
showSimpleDT( Sys.info())
```
</center>

### R
```{r rmd_runInfo_r}
pander( sessionInfo());
```

<!-- Ending the tabset section -->
## {.toc-ignore}
<br /><br />




## Load data
```{r rmd_loadData}
<<loadData>>
<<plotsPreparation>>
```
<br /><br />  




## Dimensionality reduction {.tabset .tabset-fade}

### Clusters {.tabset .tabset-fade .tabset-pills}
<!-- #### Interactive -->
<!-- ```{r rmd_plotDimReducInteractive_colorClusters} -->
<!-- dimReducWidth  = 950; -->
<!-- dimReducHeight = 800; -->
<!-- initialPointSize = 5; -->
<!-- showDimReducClusterLabels = FALSE; -->
<!-- showLegend = TRUE; -->
<!-- <<plotDimReducInteractive_colorClusters>> -->
<!-- rm(showDimReducClusterLabels); -->
<!-- rm(showLegend); -->
<!-- ``` -->
<!-- <br />   -->

#### Raster
```{r rmd_plotDimReducRaster_colorClusters}
showDimReducClusterLabels = TRUE;
showLegend = FALSE;
<<plotDimReducRaster_colorClusters>>
rm(showDimReducClusterLabels);
rm(showLegend);
```
<br />  

#### Raster (no labels)
```{r rmd_plotDimReducRaster_colorClusters_noLabels}
showDimReducClusterLabels = FALSE;
showLegend = FALSE;
<<plotDimReducRaster_colorClusters>>
rm(showDimReducClusterLabels);
rm(showLegend);
```
<br />  

### Clusters + HTOs {.tabset .tabset-fade .tabset-pills}
<!-- #### Interactive -->
<!-- ```{r rmd_plotDimReducInteractive_colorClusters_symbolHTOs} -->
<!-- dimReducWidth  = 950; -->
<!-- dimReducHeight = 800; -->
<!-- initialPointSize = 5; -->
<!-- showDimReducClusterLabels = FALSE; -->
<!-- showLegend = TRUE; -->
<!-- <<plotDimReducInteractive_colorClusters_symbolHTOs>> -->
<!-- rm(showDimReducClusterLabels); -->
<!-- rm(showLegend); -->
<!-- ``` -->
<!-- <br />   -->

<!-- #### Raster -->
```{r rmd_plotDimReducRaster_colorClusters_symbolHTOs, fig.dim = c(11, 9)}
showDimReducClusterLabels = FALSE;
showLegend = TRUE;
<<plotDimReducRaster_colorClusters_symbolHTOs>>
rm(showDimReducClusterLabels);
rm(showLegend);
```
<br />  

### Facet HTOs {.tabset .tabset-fade .tabset-pills}
<!-- #### Interactive -->
<!-- ```{r rmd_plotDimReducInteractiveGG_colorClusters_facetHTOs, warning = FALSE} -->
<!-- dimReducWidth  = 950; -->
<!-- dimReducHeight = 800; -->
<!-- initialPointSize = 5; -->
<!-- showDimReducClusterLabels = FALSE; -->
<!-- showLegend = TRUE; -->
<!-- <<plotDimReducInteractiveGG_colorClusters_facetHTOs>> -->
<!-- rm(showDimReducClusterLabels); -->
<!-- rm(showLegend); -->
<!-- ``` -->
<!-- <br />   -->

<!-- #### Raster -->
```{r rmd_plotDimReducRaster_colorClusters_facetHTOs, fig.dim = c(11, 9)}
showDimReducClusterLabels = FALSE;
showLegend = TRUE;
<<plotDimReducRaster_colorClusters_facetHTOs>>
rm(showDimReducClusterLabels);
rm(showLegend);
```
<br />  

### Facet Clusters {.tabset .tabset-fade .tabset-pills}
<!-- #### Interactive -->
<!-- ```{r rmd_plotDimReducInteractiveGG_colorHTOs_facetClusters, warning = FALSE} -->
<!-- dimReducWidth  = 950; -->
<!-- dimReducHeight = 800; -->
<!-- initialPointSize = 5; -->
<!-- showDimReducClusterLabels = FALSE; -->
<!-- showLegend = TRUE; -->
<!-- <<plotDimReducInteractiveGG_colorHTOs_facetClusters>> -->
<!-- rm(showDimReducClusterLabels); -->
<!-- rm(showLegend); -->
<!-- ``` -->
<!-- <br />   -->

<!-- #### Raster -->
```{r rmd_plotDimReducRaster_colorHTOs_facetClusters, fig.dim = c(11, 9)}
showDimReducClusterLabels = FALSE;
showLegend = TRUE;
<<plotDimReducRaster_colorHTOs_facetClusters>>
rm(showDimReducClusterLabels);
rm(showLegend);
```
<br />  




## Monitored genes
```{r rmd_heterogeneity_monitoredGenes, warning = TRUE}
<<heterogeneity_monitoredGenes>>
```

### Monitored genes heatmap

Expression heatmap of monitored genes sorted by Seurat clusters:

```{r rmd_heterogeneity_monitoredGenes_heatmap, fig.dim = c(8, min(1+(length(unlist(MONITORED_GENES))/10), 150))}
<<heterogeneity_monitoredGenes_heatmap>>
```

### Monitored genes expression {.tabset .tabset-fade}

#### UMAP {.tabset .tabset-fade}
```{r rmd_heterogeneity_monitoredGenes_expression_projection_UMAP, fig.dim = c(5, 5), out.width='25%'}
useReduction='umap'

cat("\n<a href='#clusters-projection'>");
<<heterogeneity_dimReduc_thumbnail>>
cat("</a> \n \n");

<<heterogeneity_monitoredGenes_expression_projection>>

cat(" \n \n");

rm("useReduction");
```

<!-- Ending the tabset section -->
#### {.toc-ignore}

#### UMAP by HTO {.tabset .tabset-fade}
```{r rmd_heterogeneity_monitoredGenes_expression_projection_UMAP_split_factorHTO, fig.dim = c(18, 4.5)}
useReduction='umap'
metadataColSplit='factorHTO'

cat("\n<a href='#clusters-projection'>");
<<heterogeneity_dimReduc_thumbnail>>
cat("</a> \n \n");

<<heterogeneity_monitoredGenes_expression_projection_splitByMetadata>>

cat(" \n \n");

rm("metadataColSplit");
rm("useReduction");
```

<!-- Ending the tabset section -->
#### {.toc-ignore}

<!-- #### tSNE {.tabset .tabset-fade} -->
<!-- ```{r rmd_heterogeneity_monitoredGenes_expression_projection_tSNE, fig.dim = c(4, 4), out.width='25%'} -->
<!-- useReduction='tsne' -->

<!-- cat("\n<a href='#clusters-projection'>"); -->
<!-- #<<heterogeneity_dimReduc_thumbnail>> -->
<!-- cat("</a> \n \n"); -->

<!-- #<<heterogeneity_monitoredGenes_expression_projection>> -->

<!-- rm("useReduction"); -->
<!-- ``` -->

<!-- <!-- Ending the tabset section --> -->
<!-- #### {.toc-ignore} -->

#### Clusters violin {.tabset .tabset-fade}
```{r rmd_heterogeneity_monitoredGenes_expression_violin, fig.dim = c(4.5, 4.5), out.width='25%'}
cat("\n<a href='#clusters-projection'>");
useReduction='umap'; #c( 'umap', 'tsne');
<<heterogeneity_dimReduc_thumbnail>>
rm("useReduction");
cat("</a> \n \n");

<<heterogeneity_monitoredGenes_expression_violin>>

cat(" \n \n");
```

<!-- Ending the tabset section -->
#### {.toc-ignore}

#### Clusters violin by HTO {.tabset .tabset-fade}
```{r rmd_heterogeneity_monitoredGenes_expression_violin_split_factorHTO, fig.dim = c(4.5, 13.5), out.width='25%'}
cat("\n<a href='#clusters-projection'>");
useReduction='umap'; #c( 'umap', 'tsne');
<<heterogeneity_dimReduc_thumbnail>>
rm("useReduction");
cat("</a> \n \n");

metadataColSplit='factorHTO'
<<heterogeneity_monitoredGenes_expression_violin_splitByMetadata>>

cat(" \n \n");

rm("metadataColSplit")
```

<!-- Ending the tabset section -->
#### {.toc-ignore}

#### Clusters expression {.tabset .tabset-fade}
```{r rmdheterogeneity_monitoredGenes_expressionCircles_by_cluster, fig.dim = c(6, 5)}

<<heterogeneity_monitoredGenes_expressionCircles_by_cluster>>

cat(" \n \n");

```

<!-- Ending the tabset section -->
#### {.toc-ignore}


#### Clusters expression by HTO {.tabset .tabset-fade}
```{r rmd_heterogeneity_monitoredGenes_expressionCircles_by_cluster_splitByMetadata, fig.dim = c(5, 15)}

metadataColSplit='factorHTO'
<<heterogeneity_monitoredGenes_expressionCircles_by_cluster_splitByMetadata>>

cat(" \n \n");

rm("metadataColSplit")
```

<!-- Ending the tabset section -->
#### {.toc-ignore}








<!-- Ending the tabset section -->
### {.toc-ignore}
<br /><br />




## Modules analysis
```{r rmd_heterogeneity_modules}
<<heterogeneity_modules>>
<<heterogeneity_modules_scoring>>
```

### Modules scoring heatmap

Heatmap of modules scoring sorted by Seurat clusters:

```{r rmd_heterogeneity_modules_heatmap, fig.dim = c(8, min(1+(length(MODULES_GENES)/10), 150))}
<<heterogeneity_modules_heatmap>>
```

### Modules scoring {.tabset .tabset-fade}

#### UMAP {.tabset .tabset-fade .tabset-pills}
```{r rmd_heterogeneity_modules_expression_projection_UMAP, fig.dim = c(6, 6), out.width='25%'}
useReduction='umap'

cat("\n<a href='#clusters-projection'>");
#<<heterogeneity_dimReduc_thumbnail>>
cat("</a> \n \n");

#<<heterogeneity_modules_expression_projection>>
#<<heterogeneity_modules_expression_projection_pngFile>>

rm("useReduction");
```

<!-- Ending the tabset section -->
#### {.toc-ignore}

<!-- #### tSNE {.tabset .tabset-fade .tabset-pills} -->
<!-- ```{r rmd_heterogeneity_modules_expression_projection_tSNE, fig.dim = c(6, 6), out.width='25%'} -->
<!-- useReduction='tsne' -->

<!-- cat("\n<a href='#clusters-projection'>"); -->
<!-- #<<heterogeneity_dimReduc_thumbnail>> -->
<!-- cat("</a> \n \n"); -->

<!-- #<<heterogeneity_modules_expression_projection>> -->
<!-- #<<heterogeneity_modules_expression_projection_pngFile>> -->

<!-- rm("useReduction"); -->
<!-- ``` -->

<!-- <!-- Ending the tabset section --> -->
<!-- #### {.toc-ignore} -->

#### Clusters violin {.tabset .tabset-fade}
```{r rmd_heterogeneity_modules_expression_violin, fig.dim = c(6, 6), out.width='25%'}
cat("\n<a href='#clusters-projection'>");
useReduction='umap'; #c( 'umap', 'tsne');
#<<heterogeneity_dimReduc_thumbnail>>
rm("useReduction");
cat("</a> \n \n");

#<<heterogeneity_modules_expression_violin>>
#<<heterogeneity_modules_expression_violin_pngFile>>
```

<!-- Ending the tabset section -->
#### {.toc-ignore}




```{r rmd_final_saveSessionImage}
#<<final_saveSessionImage>>
```




